Galaxy Zoo: Morphological Classifications for 120,000 Galaxies in HST Legacy Imaging

### 2016-09-06 writeup/gz_hubble_data_draft_2.tex
adding 2nd draft file of writeup

@@

\noindent where $\hat\zeta$ is the correction factor applied to each galaxy as
a function of surface brightness. The best-fit parameters to the linear fit
(from least-squares optimization) are $\zeta_0=0.50$, $\zeta_1=-0.03$. To make the
final debiased correction, the simple exponential form of
Equation~\ref{eqn:fzeta} is modified to bound the debiased vote fractions
between $f$ and 1:

\begin{equation}
f_\textrm{features,debiased} = 1 - (1 - f_\mathrm{features,weighted})e^{\frac{z-z_0}{\hat\zeta}}
\label{eqn:fzeta_mod}
\end{equation}

\noindent where $f_\mathrm{features,weighted}$ is the weighted vote fraction
described in Section~\ref{ssec:weighting}.

\subsection{Effects of morphological debiasing}\label{ssec:zeta_results}

The objective is to use these data to predict, for a galaxy with a measured
\ffeaturesz{} value, what its \ffeatures{} value \emph{would have been} if it
had been viewed at $z=0.3$. This predicted value is defined as the ``debiased''
vote fraction \ffeaturesdebiased, and is calculated by applying a correction to
the measured value of \ffeatures, determined by the $\zeta$ function described
in the previous section. A reliable predicted value can be obtained so long as
the relationship between \ffeaturesz{} and \ffeaturesrest{} is single-valued;
that is, for a given \ffeaturesz, there is exactly one corresponding value of
\ffeatures{} at $z=0.3$.

The unshaded regions in Figure~\ref{fig:f_vs_f} define discrete ranges of redshift,
surface brightness, and \ffeatures{} within which a galaxy must lie in order for the
debiased correction to be confidently applied.
While the appropriate correctable regions were defined as discrete bins, the true
correctable region is assumed to be a smooth function of $z$, $\mu$, and
\ffeatures{}. To define this smooth space, we calculate the shape of the convex
hull that encloses the correctable and lower-limit \ferengi{} galaxies in
$z$-$\mu$-\ffeatures{} space. The boundaries are then adjusted until the
contamination from both groups is minimised. The resulting hulls define the
correctable and lower-limit regions for categorising the \hst{} galaxies. The
results of this method and final categorisation of the \hst{} sample are in
Table~\ref{tbl:hubble_debiasable}. Of the galaxies at redshift higher than
$z=0.3$, 17\% can be debiased using the $\zeta$ method, 27\% cannot be debiased
since the relationship between \ffeaturesz{} and \ffeaturesrest{} is not
monotonic, and 56\% cannot be debiased since they either have an unknown
redshift or insufficient numbers of \ferengi{} images in their $z$-$\mu$ bin
to determine a reliable value of $\zeta$.

For the ``lower-limit'' galaxies for which a single debiased \ffeatures{} value
cannot be confidently assigned, the \emph{range} of debiased values is
estimated using a method visualised in Figure~\ref{fig:f_vs_f_zoom}. This uses the
\ferengi{} simulated data to analyse the range of intrinsic \ffeaturesrest{}
values for any given observed \ffeatures{} value, again as a function of
surface brightness and redshift. In each unclassifiable (shaded blue) $z$,$\mu$ bin, 
we compute the spread of
intrinsic values of \ffeaturesrest{} for five quantiles of observed \ffeatures{}
(corresponding to the ``clean'' thresholds used in prior GZ publications).
The range of intrinsic values for GZH is defined as the inner 80\% of the data,
represented by the orange bars in Figure~\ref{fig:f_vs_f_zoom}. For any galaxy which
cannot be directly debiased by the $\zeta$ method, these ranges are used to
denote the upper and lower limits on the expected values \ffeaturesrest{} as a
function of the observed \ffeatures.

@@

In the \ferengi{} data, these numbers severely limit the amount of information
that can be extracted for the higher-tier questions. The debiasing technique
used (Section~\ref{ssec:zeta}) requires that at least 10~classifiers answer each
question for a galaxy with $z_\mathrm{sim}=0.3$ \emph{and} the corresponding
image at higher redshift. This requirement is (by default) met by all galaxies
for the smooth/features question. However, this is often \emph{not} met for
questions beyond Task~01. On average, $60\%\pm24\%$ of the galaxies do not have
sufficient data to measure a correction, as compared to 2.0\% achieved for
Task~01 (Table~\ref{tbl:ferengi_corrections}). Of the remaining
\ferengi{} galaxies which did meet this requirement, polynomials were fit to
the data in each redshift-surface brightness bin, as was done for Task~01. The
goodness-of-fit was evaluated in each bin, for each Task, using a normalised
$\chi^2$ metric. The mean $\chi^2$ for all higher-order tasks was
$0.10\pm0.04$, significantly larger than the Task~01 value of $0.04$. Based on
this statistic, the data cannot not be modelled with sufficient confidence for
the higher order tasks. An example of the debiasing technique applied to
\fbar{} is shown in Section~\ref{sec:ferengi_bar}; visual inspection of
Figure~\ref{fig:f_vs_fbar} confirms that the points in each bin are scattered
to a degree at which no strong correlation is present.  For these reasons,
debiased vote fractions are only offered in the GZH catalogue for Task~01
(smooth/features).

@@

We repeated all analyses comparing morphologies between GZH and external
publications on the entire GZH dataset, rather than the volume-limited sample
($z<1.0,m<22.5$) discussed above. If galaxies at all redshifts and at the
original magnitude limit of $m<23.5$ are used, the correlations between all of
the automatic morphological populations and GZH decrease significantly.  The
exceptions are the GEMS galaxies, which are the only sample that was visually
classified and had a shallower magnitude cut on their catalogue.  Based on
these comparisons, we suggest that use of morphological catalogues with these
\hst{} images be limited to galaxies with $m_\mathrm{I|i|z}<22.5$; fainter
targets have been demonstrated to be less reliable, regardless of the method of
morphological classification being used.

@@

\subsection{Debiasing higher-order tasks: \fbar}
\label{sec:ferengi_bar}

To determine the rate at which the vote fraction for any Task decreases with
redshift or surface brightness, data from simulated \ferengi{} images were
modelled with functions in discrete redshift and surface brightness bins
(Section~\ref{ssec:zeta_results}). A cut of $N>10$ was placed on the number of
votes to reduce the error in the vote fractions and ensure they were
well-sampled before applying a fit to the data. This requirement was only met
for Task~01, which significantly reduced the goodness-of-fit for the model
functions as compared to the smooth/features task. Here we show an
example of results obtained for the bar Task. Figure~\ref{fig:f_vs_fbar}
shows that both the low sample rate and the degree of scatter in each bin
inhibit fitting a reliable function that predicts $f_{{\rm bar},z=0.3}$.
Table~\ref{tbl:ferengi_bar} summarises the correction types for the
\ferengi{} data; 71\% of the galaxies simulated did not meet the GZH
requirements for well-sampled vote fractions for $f_\mathrm{bar}$, as
compared to 2\% for $f_\mathrm{features}$. Similar results were obtained
for all higher-order Tasks in GZH (Section~\ref{ssec:higher_order_tasks}).

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/p_vs_p_SB_redshift_bar.pdf}
\caption{Similar to Figure~\ref{fig:f_vs_f}, showing effects of redshift bias
for 3,449~images in the \ferengi{} sample. Each point in a given redshift
and surface brightness bin represents a unique galaxy. The $y$-axis in each
bin is the \ffeatures{} value of the image of that galaxy redshifted to the
value corresponding to that redshift bin. The $x$-axis is the \ffeatures{}
value of the image of the same galaxy redshifted to $z=0.3$. The dashed
black lines represent the best-fit polynomials to the data in each square.
The solid black line represents \ffeaturesz=\ffeaturesrest. Regions in
which there is a single-valued relationship between \ffeatures{} at high
redshift and at $z=0.3$ are marked in white; those in which there is not
are blue, and those with not enough data ($N<5$) are grey. A larger version of
the bin outlined at $z=1.0$ and $20.3 < \mu < 21.0$ $\rm (mag/arcsec^2)$ is
shown in Figure~\ref{fig:f_vs_fbar_zoom}. The mean normalised $\chi^2$ of the
change in vote fraction is 0.08. None of the correlations for higher-order
tasks (including bars) were applied to galaxies in the GZH catalogue.}
\label{fig:f_vs_fbar}
\end{figure*}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/z1_mu20_subplot2_bar.pdf}
\caption{A larger version of the dark-outlined square in Figure~\ref{fig:f_vs_fbar}, containing
\ferengi{} galaxies that have been artificially redshifted to $z=1.0$ and have
surface brightnesses between $20.3 < \mu < 21.0$ $\rm (mag/arcsec^2)$. The orange bars represent the inner
80$^\mathrm{th}$ percentiles of the data, the boundaries of which are the lower
and upper limits of the debiased values.}
\label{fig:f_vs_fbar_zoom}
\end{figure}

\begin{table}ー
\label{tbl:ferengi_bar}
\caption{Morphological correction categories for the \ferengi{} images summarised in
Figure~\ref{fig:f_vs_fbar}. Correctable images had a single-valued relationship
between their measured \fbar{} values at high and low redshifts (white regions
in Figure~\ref{fig:f_vs_fbar}). Galaxies with a lower-limit on \fbar{} had a
non single-valued relationship (blue regions). NEI images had undetermined
relationships due to a lack of data ($N<5$) in their corresponding $z$-$\mu$
bins (grey regions). Only 17\% (maximum) of \ferengi{} galaxies in the sample
were considered ``correctable'', which is not sufficient to compute a $\zeta$
function applicable to the \hst{} data.   \label{tbl:ferengi_bar_corrections}}
\begin{tabular}{lrr}
\hline \hline
                                   & $N$    & \%        \\
\hline 
Correctable                        & 664    & 17\%      \\
Lower-limit                        & 483    & 12\%      \\
NEI                                & 2,803  & 71\%      \\
\hline
Total                              & 3,950  & 100\%     \\
\hline \hline
\end{tabular}
\end{table}

>>>>

\subsection{Effects of morphological debiasing}\label{ssec:zeta_results}
The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the galaxy.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. The combination of all
such parameters forms a high-dimensional space, and it is not clear how to
separate this into individual effects. Instead, the method used here employs
only two parameters as a simple model of the effect of redshift brightness on classification: the
surface brightness ($\mu$; intrinsic) and redshift ($z$;
extrinsic).

The objective is to use these data to predict, for a galaxy with a measured
\ffeaturesz{} value, what its \ffeatures{} value \emph{would have been} if it
had been viewed at $z=0.3$. This predicted value is defined as the ``debiased''
vote fraction \ffeaturesdebiased, and is calculated by applying a correction to
the measured value of \ffeatures, determined by the $\zeta$ function described
in the following section. A reliable predicted value can be obtained so long as
the relationship between \ffeaturesz{} and \ffeaturesrest{} is single-valued;
that is, for a given \ffeaturesz, there is exactly one corresponding value of
\ffeatures{} at $z=0.3$.

The unshaded regions in Figure~\ref{fig:f_vs_f} define discrete ranges of redshift,
surface brightness, and \ffeatures{} within which a galaxy must lie in order for the
debiased correction to be confidently applied.
While the appropriate correctable regions were defined as discrete bins, the true
correctable region is assumed to be a smooth function of $z$, $\mu$, and
\ffeatures{}. To define this smooth space, we calculate the shape of the convex
hull that encloses the correctable and lower-limit \ferengi{} galaxies in
$z$-$\mu$-\ffeatures{} space. The boundaries are then adjusted until the
contamination from both groups is minimised. The resulting hulls define the
correctable and lower-limit regions for categorising the \hst{} galaxies. The
results of this method and final categorisation of the \hst{} sample are in
Table~\ref{tbl:hubble_debiasable}. Of the galaxies at redshift higher than
$z=0.3$, 17\% can be debiased using the $\zeta$ method, 27\% cannot be debiased
since the relationship between \ffeaturesz{} and \ffeaturesrest{} is not
monotonic, and 56\% cannot be debiased since they either have an unknown
redshift or insufficient numbers of \ferengi{} images in their $z$-$\mu$ bin
to determine a reliable correction term ($\zeta$, described in section ~\ref{ssec:zeta}).

For the ``lower-limit'' galaxies for which a single debiased \ffeatures{} value
cannot be confidently assigned, the \emph{range} of debiased values is
estimated using a method visualised in Figure~\ref{fig:f_vs_f_zoom}. This uses the
\ferengi{} simulated data to analyse the range of intrinsic \ffeaturesrest{}
values for any given observed \ffeatures{} value, again as a function of
surface brightness and redshift. In each unclassifiable (shaded blue) $z$,$\mu$ bin, 
we compute the spread of
intrinsic values of \ffeaturesrest{} for five quantiles of observed \ffeatures{}
(corresponding to the ``clean'' thresholds used in prior GZ publications).
\hl{The range of intrinsic values for GZH is defined as the inner 68\% of the data, corresponding to
where the upper and lower limits of this range correspond to} $\pm{1~\sigma}$ \hl{ above and 
below the mean. This range is represented by the orange bars in Figure}~\ref{fig:f_vs_f_zoom}. For any galaxy which
cannot be directly debiased by the $\zeta$ method, these ranges are used to
denote the upper and lower limits on the expected values \ffeaturesrest{} as a
function of the observed \ffeatures.

@@

\noindent where $\hat\zeta$ is the correction factor applied to each galaxy as
a function of surface brightness. The best-fit parameters to the linear fit
(from least-squares optimization) are $\zeta_0=0.50$, $\zeta_1=-0.03$. To make the
final debiased correction, we use Equation~\ref{eqn:fzeta}, rearranged for clarity:

\begin{equation}
f_\textrm{features,debiased} = 1 - (1 - f_\mathrm{features,weighted})e^{\frac{-(z-z_0)}{\hat\zeta}}
\label{eqn:fzeta_mod}
\end{equation}

\noindent where $f_\mathrm{features,weighted}$ is the weighted vote fraction
described in Section~\ref{ssec:weighting}, and here $f_{\rm features,debiased}$ is bounded 
between $f_{\rm features,weighted}$ and 1.

@@

In the \ferengi{} data, these numbers severely limit the amount of information
that can be extracted for the higher-tier questions. The debiasing technique
used (Section~\ref{ssec:zeta}) requires that at least 10~classifiers answer each
question for a galaxy with $z_\mathrm{sim}=0.3$ \emph{and} the corresponding
image at higher redshift. This requirement is (by default) met by all galaxies
for the smooth/features question. However, this is often \emph{not} met for
questions beyond Task~01. On average, $60\%\pm24\%$ of the galaxies do not have
sufficient data to measure a correction, as compared to 2.0\% achieved for
Task~01 (Table~\ref{tbl:ferengi_corrections}).\hl{This left each surface brightness/redshift bin 
with insufficient data points to fit a function with confidence; on average only 9 points
remained in each bin (as compared to an average of 59 points per bin for Task~01). For these reasons,
debiased vote fractions are only offered in the GZH catalogue for Task~01
(smooth/features).  }

@@

We repeated all analyses comparing morphologies between GZH and external
\hl{publications} on the entire GZH dataset, rather than the volume-limited sample
($z<1.0,m<22.5$) discussed above. If galaxies at all redshifts and at the
original magnitude limit of $m<23.5$ are used, the correlations between all of
the automatic morphological populations and GZH decrease significantly.  The
exceptions are the GEMS galaxies, which are the only sample that was visually
classified and had a shallower magnitude cut on their catalogue.  Based on
these comparisons, we suggest that use of morphological catalogues with these
\hst{} images be limited to galaxies with $m_\mathrm{I|i|z}<22.5$; fainter
targets have been demonstrated to be less reliable, regardless of the method of
morphological classification being used.

@@
